AHABus \& FCORE:\\Modular Bus for High-Altitude Balloon Missions

### 2017-04-09 chapters/03-methodology.tex
Begin FCORE

The second experimental setup was much simpler: instead of using an hybrid
solution like the one described in the previous section, it relied on the
\acrshort{i2c} standard. The protocol was chosen because it provides a two-way
communication interface over two wires, for up to 127 devices. The major
difference with a \acrshort{uart} serial interface is that \acrshort{i2c} is
asymmetric: one device acts as the bus controllers, and passive devices can
only receive or transmit data when addressed by it.

@@

@@

\subsubsection{Payload Priority System}

\subsubsection{Mission Design and Configuration Files}
\label{sssec:conf-file}

@@

\subsection{Extracting Raw Binary Stream}

\subsection{Packet}

>>>>

The second experimental setup was much simpler: instead of using an hybrid
solution like the one described in the previous section, it relied on the
\acrshort{i2c} standard. The protocol was chosen because it provides a two-way
communication interface over two wires, for up to 127 devices. The major
difference with a \acrshort{uart} serial interface is that \acrshort{i2c} is
asymmetric: one device acts as the bus controller, and passive devices can
only receive or transmit data when addressed by it.

@@

%TODO: formulation
The main role of the flight software is to link the subsystems discussed
previously and handle the data flowing through the system, from each payload's
raw data to the resulting radio frames. It is also responsible for keeping track
of the platform geographic co-ordinates and its altitude.

To avoid the added workload of designing and programming a real-time firmware
from nothing, the \acrfull{fcore}, \acrshort{ahabus}'s flight software, was
based upon FreeRTOS and the SDK provided by Espressif for the ESP8266
micro-controllers, which was used as the \acrshort{obc} for the prototype.

FreeRTOS is a real-time kernel that provides a prioritised task model: each task
is given a function (usually, an event loop) to run and a priority level. At
regular intervals, the kernel establishes a list of all the tasks that must run,
and gives processor time to the one with the highest priority. Such a
deterministic system is desirable for applications like \acrshort{ahabus}, which
must run on embedded devices with limited computing power, and where certain
actions must imperatively run at regular intervals. \acrshort{fcore} was
designed around three main tasks, along with the standalone radio subsystem

\subsubsection{Payload Bus Task}
Each payload is assigned its own task, with a running
interval depending on how the payload was configured before flight. Each task
fetches data from its payload, packetises it and adds it to the radio queue.

\subsubsection{System Health Task}

The system task runs at fixed intervals to check the status of each subsystem
and create a system health message, sent to the ground station through the
radio subsystem. It was assigned the highest priority level, so that even if a
subsystem were to fail and not explicitly release control, FreeRTOS would give
the System Task the execution context, thus allowing health packets to be sent
to the ground station.

The health of each subsystem is stored globally, and each system can indicate
an issue by modifying its entry. There are four health states that subsystems
can have, represented each by a single ASCII character: booting (B), up (U),
recovering (R) and down (D). Each system is flagged as booting on startup.
Payloads are flagged as recovering if the last attempt to poll their data
failed, and down if more than five attempts to communicate failed in a row.
the \acrshort{gnss} subsystem is flagged as recovering if no valid position
fix was received from the receiver, and down if the received messages were not
parsed correctly.

Health messages were designed to be human-readable: this improves the chances
that ground teams can understand the status of the platform, even if the
message's packet and frame were to be corrupted beyond the Frame Layer's
\acrshort{fec} limits.

The message is sent as a list of comma-separated entries. Each entry starts
with a letter indicating its type (G for the \acrshort{gnss} subsystem, P for
payloads). In the case of payloads, the type is followed by the payload's
identifier as a two-character hexadecimal number. The identifier is then
followed by a forward slash (``/'') and the status character as discussed above.
Listing \ref{lst:health-message} shows an example health message.

\begin{lstlisting}[caption=FCORE - Example Health Message, label=lst:health-message]
FCORE//SYS_HEALTH G/U,P0A/R,P14/U
\end{lstlisting}

\subsubsection{GNSS Task}

The \acrshort{gnss} task was designed to run at fixed intervals to parse the
\acrshort{gnss} receiver's output, and update the platform's cached
co-ordinates and altitude. This cached location is the one included in radio
packets.

\subsubsection{Radio Subsystem}

\subsection{Mission Design and Configuration Files}
\label{sssec:conf-file}

@@

@@

\subsection{Extracting Raw Binary Stream}
\label{sec:binary-extraction}

\subsection{Packet Parser}

### 2017-04-09 chapters/04-results.tex
Start of discussion
Begin FCORE

The following section presents the results of the tests done as part of the
evaluation process of the project, as discussed in section \ref{sec:testing}.

% \section{Component Testing}
%
% \section{Integrated Testing}

@@

>>>>

The following section presents the observations and numerical results obtained
during the tests done as part of the evaluation process of the project, as
discussed in section \ref{sec:testing}.

\section{Observed Issues}

\subsection{Radio Communications}

The initial tests of the radio encoder and transmission stack ported to the
ESP8266 revealed issues with the micro-controller's interrupt system: events for
both \acrshort{uart} share the same interrupt, and \texttt{UART0} events (used
when communicating with the \acrshort{gnss} receiver) were detected at the same
time as \texttt{UART1} events. Additionally, the design of the ESP8266 meant
that the interrupts would fire not just when the transmission buffer crossed
its low threshold, but whenever it was filled below the threshold, locking the
controller into an interrupt loop and eventually triggering the watchdog timer.

Once those issues were fixed, as discussed in section \ref{sec:flight-software},
the radio protocol performed without issues on the transmission side. Reception
proved mostly stable apart from one main issue caused by the frame format: since
the end of a frame is detected by the number of byte received, byte losses
while receiving a frame would cause the start of the subsequent frame to be
counted as the end of the current one. As a result, dropped bytes in a frame
would most often cause the next frame to not be detected even if it was formed
properly.

\subsection{Payload Bus}

Though the payload bus was found to be reliable overall, two issues were found
during platform testing. First, the default \acrshort{i2c} library provided
with the ESP8266 SDK had a very short clock stretch timeout, which meant that
payloads had a very short time window to respond to any request from the bus
controller, leading to a high rate of failed communications until the stretch
value was raised. Second, there were some cases where the last command from the
bus controller (clear the connection flag register) would fail. As a result,
the payload would be prevented to write any new data until the next
communication from the bus controller.

@@

Figure \ref{fig:ditl-bandwidth} shows that the average data reception rate was
stable over the whole simulation, with an average of \SI{11}{Bps}, or
\SI{121}{bps} (out of a maximum bandwidth of \SI{200}{bps}). Over the full
simulation's duration, an average of \SI{0.63}{Bps} were lost due to
transmission errors.
